Simulating High-Dimensional Multivariate Data

### 2021-03-01 060-conclusion-discussion.Rmd
Revise 060 prior to CSDA submission

We have introduced a general-purpose high-dimensional multivariate simulation algorithm and provide a user-friendly, high-performance `R` package named [`bigsimr`](https://schisslergroup.github.io/bigsimr/). The random vector generation method is largely inspired by NORTA (@Cario1997) and Gaussian copula-based approaches (@MB13, @BF17, @Xia17). The major contributions of this work are high-dimensional scalability, flexible modeling of dependency, and an high-performance implementation with broad potential data analytic applications for modern, big-data statistical computing.

Researchers working in multivariate computation frequently encounter such difficulties and need to find a close positive (semi-)definite matrix. A widely-available routine for this task in `R` is `matrix::nearPD`, though it is not suitable for high dimensions. To overcome this issue, we introduce `cor_nearPD` --- an implementation of @QS2006's method that computes the nearest correlation matrix to a given symmetric matrix in the Frobenius norm. This routine as many applications beyond our primary goal of random vector generation.

It is customary to compare new tools and algorithms directly to existing competing methods and software. In this study, however, we only employ our proposed methodology, as our previous work has shown that existing tools are simply not designed nor feasible to meet our high-dimensional goal (see @Li2019gpu for evaluations of the `R` `copula` package and others). For the bivariate simulations, existing packages such as [`nortaRA`](https://github.com/cran/NORTARA/blob/master/inst/doc/NORTARA.R) work well to match Pearson correlations exactly.

@@

There are limitations to the methodology and implementation. The most obvious missing feature of the proposed methodology is the inability to match a Pearson correlation matrix exactly. As discussed in the [Algorithms](algorithms) section and extensively by @XZ19, this is a computationally intense procedure and Pearson's correlation is not a natural choice to describe dependency for non-normal marginals. While we do not provide an implementation directly supporting Pearson matching, users may supply their own input Pearson correlation after employing a supplementary matching scheme [@Cario1997; @XZ19].

>>>>

We have introduced a general-purpose high-dimensional multivariate simulation algorithm and provide a user-friendly, high-performance `R` package [`bigsimr`]. The random vector generation method is inspired by NORTA [@Cario1997] and Gaussian copula-based approaches [@MB13, @BF17, @Xia17]. The major contributions of this work are methods and software for flexible, scalable simulation of HD multivariate probability distributions with broad potential data analytic applications for modern, big-data statistical computing. For example, one could simulate high-resolution time series data, such as those consistent with an auto-regressive moving average model exhibiting a specified Spearman structure. Or our methods could be used to simulate sparsely correlated data, as many HD methods assume, via specifying a *spiked correlation matrix*.

Researchers working in multivariate computation frequently encounter such difficulties and need to find a close or nearest correlation matrix. A widely-available routine for this task in `R` is `matrix::nearPD`, though it is not suitable for high dimensions. To overcome this issue, we introduce `cor_nearPD` --- an implementation of @QS2006's method that computes the nearest correlation matrix to a given symmetric matrix in the Frobenius norm. This routine as many applications beyond our primary goal of random vector generation.

It is customary to compare new tools and algorithms directly to existing competing methods and software. In this study, however, we only employ our proposed methodology, as our previous work has shown that existing `R` tools are simply not designed to meet our high-dimensional goal (see @Li2019gpu for evaluations of the `R` `copula` package and others). For the bivariate simulations, existing packages such as [`nortaRA`](https://github.com/cran/NORTARA/blob/master/inst/doc/NORTARA.R) work well to match Pearson correlations exactly.

@@

There are limitations to the methodology and implementation. We could only investigate selected multivariate distributions in our Monte Carlo studies. There may be instances that the methods do not perform well. Along those lines, we expect that correlation values close to the boundary of the feasible region could result in algorithm failure. Another issue is that for discrete distributions, we use continuous approximations when the support set is large. This could limit the scalability/accuracy for particular joint distributions.

<!-- 
The most obvious missing feature of the proposed methodology is the inability to match a Pearson correlation matrix exactly. As discussed in the [Algorithms](algorithms) section and extensively by @XZ19, this is a computationally intense procedure and Pearson's correlation is not a natural choice to describe dependency for non-normal marginals. While we do not provide an implementation directly supporting Pearson matching, users may supply their own input Pearson correlation after employing a supplementary matching scheme [@Cario1997; @XZ19].
-->
